---
title: "Clean Salty Soil Data"
author: "Irfan Ainuddin"
date: "4/6/2020"
output: html_document
---
```{r}
library(tidyverse)
```

```{r}
sd <- readRDS(file = "./data/clean_soil_data.rds")
```


```{r}
hist(sd$Year)
```
Most of the data was collected 2000-2005. 

```{r}
ggplot(sd, aes(x=sd$Year, y=sd$Mg_Sat_pct, col=as.factor(sd$Depth_1))) + 
            geom_point() + theme_bw() + theme(legend.position="top") + 
            geom_smooth(se=FALSE, method="lm") + 
            geom_smooth(aes(x=sd$Year, y=sd$Mg_Sat_pct), col="blue", se=FALSE, method='lm')
```
Overall magnesium trend is decreasing at all depths.

```{r}
#
ggplot(sd, aes(x=sd$Year, y=sd$Na_Sat_pct, col=as.factor(sd$Depth_1))) + 
            geom_point() + theme_bw() + theme(legend.position="top") + 
            geom_smooth(se=FALSE, method="lm") + 
            geom_smooth(aes(x=sd$Year, y=sd$Na_Sat_pct), col="blue", se=FALSE, method='lm') 
## irf: geom_smooth y changed from Mg_sat_pct  > Na_sat_pct
```
Overall Sodium (Na) saturation percent stays roughly the same over time. Sodium decreases by roughly 5 percent in the 0-12" depth class from 17 % to 11%  over 15 years. Sodium increases very slightly in the 13-24" depth class by 1% from 17% to 18%. Sodium increases by 5% total in the 25-36" depth class from 17% to 23%. Sodium increases by 8% total in the 37-48" depth class from 17% to 24%.

These trends suggest that sodium is leeching out of the top soil into the subsoil primarily in the two lower depth classes. 


```{r}
## lets facet_wrap() by irrigation season
ggplot(sd, aes(x=sd$Year, y=sd$Na_Sat_pct, col=as.factor(sd$Depth_1))) + 
            geom_point() + theme_bw() + theme(legend.position="top") + 
            geom_smooth(se=FALSE, method="lm") + 
            geom_smooth(aes(x=sd$Year, y=sd$Na_Sat_pct), col="blue", se=FALSE, method='lm') + 
            facet_wrap(sd$Irr_season) +
            xlab("Sampling Date") +
            ylab("Sodium (Na) saturation percent (%)") +
            title("Sodium over time by irrigation season")
```

In all irrigation seasons the sodium saturation percent is decreasing for the 0-12" depth class. The samples collected 'Before' the irrigation are most similar to the overall changes shown. When we look at changes in sodium levels after irrigation and during the winter we see a decreasing trend for all of the soil depth classes, which is reflecting the soluble nature of sodium moving or leaching out with water in the natural environment. 


Slightly higher Na Sat Pct in deeper layers. 

```{r}
ggplot(sd, aes(x=sd$Depth_1, y=sd$Na_Sat_pct, fill=sd$Depth_1)) + geom_boxplot() 
```






