---
title: "Clean Salty Soil Data"
author: "Irfan Ainuddin"
date: "4/6/2020"
output: html_document
---
```{r}
library(tidyverse)
```

```{r}
sd <- readRDS(file = "./data/clean_soil_data.rds")
```


```{r}
hist(sd$Year)
```
Most of the data was collected 2000-2005. 

```{r}
ggplot(sd, aes(x=sd$Year, y=sd$Mg_Sat_pct, col=as.factor(sd$Depth_1))) + 
            geom_point() + theme_bw() + theme(legend.position="top") + 
            geom_smooth(se=FALSE, method="lm") + 
            geom_smooth(aes(x=sd$Year, y=sd$Mg_Sat_pct), col="blue", se=FALSE, method='lm')
```
Overall magnesium trend is decreasing at all depths.

```{r}
#
ggplot(sd, aes(x=sd$Year, y=sd$Na_Sat_pct, col=as.factor(sd$Depth_1))) + 
            geom_point() + theme_bw() + theme(legend.position="top") + 
            geom_smooth(se=FALSE, method="lm") + 
            geom_smooth(aes(x=sd$Year, y=sd$Na_Sat_pct), col="blue", se=FALSE, method='lm') 
## irf: geom_smooth y changed from Mg_sat_pct  > Na_sat_pct
```
Overall Sodium (Na) saturation percent stays roughly the same over time. Sodium decreases by roughly 5 percent in the 0-12" depth class from 17 % to 11%  over 15 years. Sodium increases very slightly in the 13-24" depth class by 1% from 17% to 18%. Sodium increases by 5% total in the 25-36" depth class from 17% to 23%. Sodium increases by 8% total in the 37-48" depth class from 17% to 24%.

These trends suggest that sodium is leeching out of the top soil into the subsoil primarily in the two lower depth classes. 


```{r}
## lets facet_wrap() by irrigation season
ggplot(sd, aes(x=sd$Year, y=sd$Na_Sat_pct, col=as.factor(sd$Depth_1))) + 
            geom_point() + theme_bw() + theme(legend.position="top") + 
            geom_smooth(se=FALSE, method="lm") + 
            geom_smooth(aes(x=sd$Year, y=sd$Na_Sat_pct), col="blue", se=FALSE, method='lm') + 
            facet_wrap(sd$Irr_season)
```




Slightly higher Na Sat Pct in deeper layers. 

```{r}
ggplot(sd, aes(x=sd$Depth_1, y=sd$Na_Sat_pct, fill=sd$Depth_1)) + geom_boxplot() 
```






