---
title: "cleanining_protocols"
author: "Irfan Ainuddin"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The function do the following:

Clean Data from NA’s and Blanks
Separate the clean data – Integer dataframe, Double dataframe, Factor dataframe, Numeric dataframe, and Factor and Numeric dataframe.
View the new dataframes
Create a view of the summary and describe from the clean data.
Create histograms of the data frames.
Save all the objects
This will happen in seconds.

Package
First, load Hmisc package. I always save the original file.
The code below is the engine that cleans the data file.



```{r}
cleandata <- dataname[complete.cases(dataname),]
```


```{r}
cleanme <- function(dataname){
  
  #SAVE THE ORIGINAL FILE
  oldfile <- write.csv(dataname, file = "oldfile.csv", row.names = FALSE, na = "")
  
  #CLEAN THE FILE. SAVE THE CLEAN. IMPORT THE CLEAN FILE. CHANGE THE TO A DATAFRAME.
  #cleandata <- dataname[complete.cases(dataname),]
  cleandata <- dataname #see what happens when we dont have complete cases.
  cleanfile <- write.csv(cleandata, file = "cleanfile.csv", row.names = FALSE, na = "")
  cleanfileread <- read.csv(file = "cleanfile.csv")
  cleanfiledata <- as.data.frame(cleanfileread)
  
  #SUBSETTING THE DATA TO TYPES
  ## subset: Integer
  logicmeint <- cleanfiledata[,sapply(cleanfiledata,is.integer)] 
  ## subset: Double
  logicmedouble <- cleanfiledata[,sapply(cleanfiledata,is.double)]
  ## subset: Factor
  logicmefactor <- cleanfiledata[,sapply(cleanfiledata,is.factor)]
  ## subset: Numeric
  logicmenum <- cleanfiledata[,sapply(cleanfiledata,is.numeric)]
  
  #subset: Factor & Numeric
  mainlogicmefactors <- cleanfiledata[,sapply(cleanfiledata,is.factor) | sapply(cleanfiledata,is.numeric)]

  #VIEW ALL FILES
  View(cleanfiledata)
  View(logicmeint)
  View(logicmedouble)
  View(logicmefactor)
  View(logicmenum)
  View(mainlogicmefactors)
  
  #describeFast(mainlogicmefactors)
  
  #ANALYTICS OF THE MAIN DATAFRAME
  cleansum <- summary(cleanfiledata)
  print(cleansum)
  
  
  cleandec <- describe(cleanfiledata)
  print(cleandec)
  
  #ANALYTICS OF THE FACTOR DATAFRAME
  factorsum <- summary(logicmefactor)
  print(factorsum)
  
  factordec <- describe(logicmefactor)
  print(factordec)
  
  #ANALYTICS OF THE NUMBER DATAFRAME
  numbersum <- summary(logicmenum)
  print(numbersum)
  
  numberdec <- describe(logicmefactor)
  print(numberdec)
  
  mainlogicmefactorsdec <- describe(mainlogicmefactors)
  print(mainlogicmefactorsdec)
  
  mainlogicmefactorssum <- describe(mainlogicmefactors)
  print(mainlogicmefactorssum)
  
  #savemenow <- saveRDS("cleanmework.rds")
  #readnow <- readRDS(savemenow)
  
  #HISTOGRAM PLOTS OF ALL TYPES
  hist(cleanfiledata)
  hist(logicmeint)
  hist(logicmedouble)
  hist(logicmefactor)
  hist(logicmenum)
  #plot(mainlogicmefactors)

  save(cleanfiledata, logicmeint, mainlogicmefactors, logicmedouble, logicmefactor, logicmenum, numberdec, numbersum, factordec, factorsum, cleandec, oldfile, cleandata, cleanfile, cleanfileread,   file = "cleanmework.RData")
}
```





```{r}
cleanme(dataname)
```


When all the data frames appear, type to load the workspace as objects.


```{r}
load("cleanmework.RData")
```
Enjoy

